---
title: "R Indego Project"
author: "Andy Kwon"
date: "2024-01-26"
output: html_document
---

```{r setup, include=FALSE}
chooseCRANmirror(graphics=FALSE, ind=1)
knitr::opts_chunk$set(echo = TRUE)
```

## Business Scenario
You are a junior data analyst working on the marketing analyst team at Indego, a bike-share
company in Philadelphia. The director of marketing believes the company’s future success
depends on maximizing the number of annual memberships. Therefore, your team wants to
understand how casual riders and annual members use Indego bikes differently. From these
insights, your team will design a new marketing strategy to convert casual riders into annual
members. But first, Indego executives must approve your recommendations, so they must be
backed up with compelling data insights and professional data visualizations.

## ASK
Two questions will guide the future marketing program:

1. How do annual members and casual riders use Indego bikes differently?
2. How can Indego use digital media to influence casual riders to become members?

## PREPARE

Dataset: https://www.rideindego.com/about/data

  -For this purposes of this analysis, I will use Q1 2023 to Q4 2023 data

```{r}  
#install packages
install.packages("tidyverse")
install.packages("lubridate")
install.packages("ggplot2")
install.packages("dplyr")
library(tidyverse)
library(lubridate)
library(ggplot2)
library(dplyr)
```

```{r}
#import data sets
q1_2023 <- read.csv("/Users/andyokwon/workspace/csv_indego/indego-trips-2023-q1.csv")
q2_2023 <- read.csv("/Users/andyokwon/workspace/csv_indego/indego-trips-2023-q2.csv")
q3_2023 <- read.csv("/Users/andyokwon/workspace/csv_indego/indego-trips-2023-q3.csv")
q4_2023 <- read.csv("/Users/andyokwon/workspace/csv_indego/indego-trips-2023-q4.csv")
```

## PROCESS
First, we need to check the column names first before merging the four datasets. 
```{r}
colnames(q1_2023)
colnames(q2_2023)
colnames(q3_2023)
colnames(q4_2023)
```

Great, all column names from the datasets match. Now, we will check the datatypes of each column to make sure that all data is formatted properly. 
```{r}
str(q1_2023)
str(q2_2023)
str(q3_2023)
str(q4_2023)
```

We need to change the bike id in datasets q2, q3, and q4 to be integers instead of char.
```{r}
suppressWarnings({
  q2_2023 <- mutate(q2_2023, bike_id = as.integer(bike_id))
  q3_2023 <- mutate(q3_2023, bike_id = as.integer(bike_id))
  q4_2023 <- mutate(q4_2023, bike_id = as.integer(bike_id))
})
```

The datasets are now ready to be merged into a single dataframe.
```{r}
indego_trips = bind_rows(q1_2023, q2_2023, q3_2023, q4_2023)
```

Next, we will add columns that list the day, month, and year of each trip. This will allow us to aggregate trip data for each day, month or year.
```{r}
indego_trips$date <- as.Date(indego_trips$start_time)
indego_trips$month <- format(as.Date(indego_trips$date), "%m")
indego_trips$day <- format(as.Date(indego_trips$date), "%d")
indego_trips$year <- format(as.Date(indego_trips$date), "%Y")
indego_trips$day_of_week <- format(as.Date(indego_trips$date), "%A")
colnames(indego_trips)
```

Before performing analysis, let's look at the final dataframe
```{r}
glimpse(indego_trips)
```

## ANALYZE
We will now perform a descriptive analysis of the data and see if we can find
any trends/patterns between daily, monthly, and yearly riders. First, we will look at some basic descriptive statistics of the duration of bike trips in minutes.
```{r}
mean(indego_trips$duration)
median(indego_trips$duration)
max(indego_trips$duration)
min(indego_trips$duration)
```

Now, we will compare day(daily pass), indego30(monthly pass), and indego 365(yearly pass) riders. 
```{r}
aggregate(indego_trips$duration ~ indego_trips$passholder_type, FUN = mean)
aggregate(indego_trips$duration ~ indego_trips$passholder_type, FUN = median)
```

